% ==== Introduction ====
\section{Introduction}
\begin{frame}{Exception Handling: definition (1)}
  \begin{block}{Definition from Wikipedia, ``Exception handling'' article}
    \begin{center}
      ``{\ttfamily\small Exception handling is the process of responding to the occurrence of exceptions \textemdash{}anomalous or exceptional conditions requiring special processing\textemdash{} during the execution of a program.  In general, an exception breaks the normal flow of execution and executes a pre-registered exception handler.}''
    \end{center}
  \end{block}
  \vfill

  \begin{itemize}
    \item \textbf{In programming languages:} an exception object is prepared, which is then said to be \inlineCode{throw}n; control flow transfers to the inner-most \inlineCode{catch} block that can handle it
  \end{itemize}

  \begin{onlyenv}<2>
    \TZoverlayLayer{}{\fontsize{32}{32}\selectfont\smile}
  \end{onlyenv}
\end{frame}

\begin{frame}{Exception Handling: definition (2)}
  \begin{columns}
    \column{0.5\textwidth}
    Roughly, for C++\ldots{}
    \begin{enumerate}
    \item \textbf{User code:} Allocate memory for exception of type $T$ and throw it.
    \item \textbf{Runtime:} Identify the proper catch block ($Catch_T$).
    \item \textbf{Runtime:} Undo (unwind) affected stack frames; clean up as needed (dtors)
    \item \textbf{Runtime:} Transfer control to $Catch_T$.
    \end{enumerate}

    \column{0.5\textwidth}
    \begin{center}
      \resizebox{0.9\columnwidth}{!}{\input{fig/call-stack}}
    \end{center}
  \end{columns}

  \begin{onlyenv}<2>
    \overlayLayer{\begin{itemize}
      \item This is \alert{ABI}-dependent!
      \item Runtime needs additional information emitted by the compiler!
    \end{itemize}}
  \end{onlyenv}
\end{frame}

\begin{frame}{Let's talk about ABIs}
  \begin{block}{Informal definition}
    An ABI (Application Binary Interface) defines the low-level (machine-code-level) interfaces\\[1ex]
    \begin{tabular}{|p{0.47\textwidth}|p{0.47\textwidth}|}
      \hline
      \textbf{ABI}                                    & \textbf{API} \\
      \hline
      \hline
      low-level (binary) interface                    & high-level (source-code) interface \\
      \hline
      E.g., data layout, calling convention, \ldots{} & E.g., function name, parameter types, return type, \ldots{}\\
      \hline
    \end{tabular}
  \end{block}
  \vfill

  Well-known C++ ABIs:\begin{itemize}
    \item MSVC, used by Microsoft Visual C++
    \item Itanium C++ ABI, likely used by everything else (incl. \texttt{clang} / \texttt{gcc})\footnote{\url{https://itanium-cxx-abi.github.io/cxx-abi/}}
  \end{itemize}

  \begin{onlyenv}<2>
    \overlayLayer{\underline{A C++ ABI usually details}\begin{itemize}
      \item Exception handling
      \item Layout of RTTI structures
      \item Function call 'lowering' (e.g. \texttt{this} argument) --- \alert{but usually no calling convention!}
      \item Name mangling
    \end{itemize}}
  \end{onlyenv}
\end{frame}

\begin{frame}{Let's talk about ABIs: Calling Conventions}
  \begin{columns}
    \column{0.66\textwidth}
    A calling convention specifies\ldots{}
    \begin{itemize}
      \item \textbf{Function arguments / return values:} register / stack~(mem)?
      \item Which CPU registers are \textbf{scratch}, \textbf{caller-saved}, \textbf{callee-saved}?
      \item And, possibly, other stack frame layout details
        \bigskip
    \end{itemize}
    \vfill
    May be specified by a complementary ABI:
    \begin{itemize}
      \item \textbf{x86\_64, UNIX-based:} System V ABI\footnote{https://refspecs.linuxbase.org/}.
      \item \textbf{x86\_64, MSFT-based:} Microsoft x64 ABI
      \item \textbf{AArch64:} AAPCS64 (AArch64 Prodedure Call Standard)\footnote{https://github.com/ARM-software/abi-aa/blob/main/aapcs64/aapcs64.rst}.
    \end{itemize}

    \column{0.33\textwidth}
    \begin{center}
      \resizebox{0.95\columnwidth}{!}{\input{fig/stack-frame-aarch64}}
    \end{center}
  \end{columns}
\end{frame}
